---
title: "Compare different ways to import SAS data set with formats into R"
output:
  pdf_document: 
    number_sections: true
  html_document: default
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = T, 
                      eval = T, # change to T if you want to re-run w/ new data
                      results = 'markup',
                      warning = F)
```

```{r}
require(haven)
```


```{r, echo=FALSE}
saspath <- 'C:/Program Files/SASHome/SASFoundation/9.4/sas.exe'
sasopts <- "-nosplash -log 'c:\\temp' -ls 80 -ps 60  -nocenter -nodate -nofmterr" # see http://bit.ly/1QB4ZTb
```

# Use haven package to read in SAS data set with the corresponding original format catalog.

## Method 1

Use the original SAS format file, sisformats.sas7bcat.

The format catalog is not read with the error, "Error in df_parse_sas_file(spec_data, spec_cat, encoding = encoding, catalog_encoding = catalog_encoding,  : Failed to parse U:/projects/Sister Study/formats/sisformats.sas7bcat: Invalid file, or file has unsupported features."

```{r, eval=F}

df1= read_sas(data_file="../../Sister Study/data/dr00224_00_02/dr00224_00_02.sas7bdat",
                 catalog_file = "../../Sister Study/formats/sisformats.sas7bcat" )

```

## Method 2

Try using a format file with UTF-8 encoding, sisformats-UTF-8.sas7bcat.

Still get the same error: "Error in df_parse_sas_file(spec_data, spec_cat, encoding = encoding, catalog_encoding = catalog_encoding,: Failed to parse U:/projects/Sister Study/formats/sisformats-UTF-8.sas7bcat: Invalid file, or file has unsupported features."


```{r, eval=F}

# read original SAS data into R: only repeat if new data

df1 = read_sas(data_file="../../Sister Study/data/dr00224_00_02/dr00224_00_02.sas7bdat",
                 catalog_file = "../../Sister Study/formats/sisformats-UTF-8.sas7bcat" )


```

If I omit the catalog I can read in the dr00224_00_02.sas7bdat and get the labels but not formats.


```{r, eval=T}

df = read_sas(data_file="../../Sister Study/data/dr00224_00_02/dr00224_00_02.sas7bdat")

# example
attr(df$AL_DrinkCat6, "label")

```


## Method 3

After some investigation, try only using the columns in the format file with non-missing values.

### SAS code to alter the format catalog

```{r, engine='sas', engine.path=saspath, engine.opts=sasopts, results='markup', echo=TRUE, message=T, warning=FALSE, eval=T}

libname libf "..\..\Sister Study\formats";
*proc contents data=libf.sisformats_data order=varnum;
*run;

data libf.sisformats_datasub; 
set libf.sisformats_data(keep=fmtname start end label min max default length); 
where fmtname not in ("BCDXDTSRC","BCSTGSR","BCSTGSRC", "BCTYPESRC", "CHEMODRUG", "LATF",
"NAANALYSISF", "NADIETRPTF", "NAOUTLBF", "STAGEF", "CENSUSDIVCREATE", "CENSUSREGCREATE");
run; * These variables have overlapping ranges?;

LIBNAME fcat_out "..\..\Sister Study\formats-subset";
PROC FORMAT CNTLIN=libf.sisformats_datasub LIBRARY=fcat_out.sisformats; 
RUN;

```

This format file, .sas7bcat, works. Assuming the haven package has problems reading in columns from a format file with no values in them.

```{r, eval=T}

df1 = read_sas(data_file="../../Sister Study/data/dr00224_00_02/dr00224_00_02.sas7bdat",
                 catalog_file = "../../Sister Study/formats-subset/sisformats.sas7bcat" )

# check
attr(df1$AL_DrinkCat6, "labels")
table(df1$AL_DrinkCat6)

```
